---
title: "Lab_05"
author: "NickyNie"
date: "9/24/2021"
output: 
  - html_document
  - github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(dplyr)
```

## Preprocessing data
```{r preprocessing, cache=TRUE}
# Download the data
met <- fread("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz")
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```
## Merging data
```{r merge}
met <- merge(
  x     = met,
  y     = stations,
  by.x  = "USAFID",
  by.y  = "USAF",
  all.x = TRUE,
  all.y = FALSE
)
```
# Question 1
```{r collapsing_by_station}
station_avg <- met[,.(
  temp = mean(temp, na.rm = TRUE),
  wind.sp = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE),
  lon_avg = mean(lon, na.rm = TRUE),
  lat_avg = mean(lat, na.rm = TRUE)
), by = USAFID]

```

identify the median per variable
```{r quantile}
medians <- station_avg[,.(
  temp_50 = quantile(temp, probs = 0.5, na.rm = TRUE),
  wind.sp_50 = quantile(wind.sp, probs = 0.5, na.rm = TRUE),
  atm.press_50 = quantile(atm.press, probs = 0.5, na.rm = TRUE)
) ]

medians
```

find stations that are the closest to these (hint: `which.min()`)
```{r median_temp}
station_avg[, temp_dist := abs(temp - medians$temp_50)]
median_temp_station <- station_avg[order(temp_dist)][1]
median_temp_station
```
The median temperature station is  `r median_temp_station$USAFID`

```{r median_wind.sp}
station_avg[, wind.sp_dist := abs(wind.sp - medians$wind.sp_50)]
median_wind.sp_station <-station_avg[order(wind.sp_dist)][1]
median_wind.sp_station 
```
The median wind speed station is  `r median_wind.sp_station$USAFID`

```{r median_atm.press}
station_avg[, atm.press_dist := abs(atm.press - medians$atm.press_50)]
median_atm.press_station <- station_avg[order(atm.press_dist)][1]
median_atm.press_station
```
The median atmospheric pressure station is  `r median_atm.press_station$USAFID`

# Question 2
recover the state variable, by MERGING
```{r}
station_avg <- merge(
  x = station_avg, 
  y = stations, 
  by.x = "USAFID", 
  by.y = "USAF", 
  all.x = TRUE, 
  all.y = FALSE
  )
```

compute the medians per state
```{r}
station_avg[, temp_50 := quantile(temp, probs = 0.5, na.rm = TRUE),by = STATE]
station_avg[, wind.sp_50 := quantile(wind.sp, probs = 0.5, na.rm = TRUE),by = STATE]
```

```{r}
station_avg[, eudist:= sqrt(
  (temp - temp_50)^2+(wind.sp - wind.sp_50)^2
)]
station_avg
```
Find the most representative station per state
```{r}
station_state <- station_avg[ , .SD[which.min(eudist)], by = STATE]
station_state
```
# Question 3
find the midpoint of each state
```{r midpoint}
mid <- met[, .(
  lon_50 = quantile(lon, probs = 0.5, na.rm = TRUE),
  lat_50 = quantile(lat, probs = 0.5, na.rm = TRUE)
), by = STATE]
```

```{r}
station_avg <- merge(
  x = station_avg,
  y = mid,
  by = "STATE"
)
```

```{r}
station_avg[, mid_eudist:= sqrt(
  (lat_avg - lat_50)^2+(lon_avg - lon_50)^2
)]
station_avg
```
```{r mid_station}
mid_station_state <- station_avg[ , .SD[which.min(mid_eudist)], by = STATE]
mid_station_state
```
merge data of median and mid stations
```{r merge_stations}
station_state[, type := "median"]
mid_station_state[, type := "mid"]
target_stations <- rbind(station_state, mid_station_state, fill = TRUE)
```

draw the map
```{r leaflet}
library(leaflet)
pal <- colorFactor(c("Blue","DarkOrange"), domain = target_stations$type)
leaflet() %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    data = target_stations,
    lat = ~lat_avg, lng = ~lon_avg, 
    opacity = 1, fillOpacity = 1, radius = 400, color = ~pal(target_stations$type)
    )
```
# Question 4
```{r}
met[, state_temp := mean(temp, na.rm = TRUE),by = STATE]
met[, state_wind.sp := mean(wind.sp, na.rm = TRUE),by = STATE]
met[, state_atm.press := mean(atm.press, na.rm = TRUE),by = STATE]
met[, temp_cat := fifelse(
  state_temp < 20, "low-temp", 
  fifelse(state_temp < 25, "mid-temp", "high-temp"))
  ]
```

make sure we do not have NAs
```{r}
table(met$temp_cat, useNA = "always")
```

```{r summarize}
tab <- met[, .(
  N_entries  = .N,
  N_stations = length(unique(USAFID)),
  avg_temp = mean(state_temp),
  avg_wind.sp = mean(state_wind.sp),
  avg_atm.press = mean(state_atm.press)
), by = temp_cat]

knitr::kable(tab)
```

