---
title: "Assignment_03"
author: "NickyNie"
date: "11/4/2021"
output: 
  - html_document
  - github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(xml2)
library(stringr)
library(dplyr)
library(tidytext)
library(tidyverse)
library(tidyr)
library(ggplot2)
```

## APIs
Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

As we did in lab 7. Create a dataset containing the following:

1. Pubmed ID number,
2. Title of the paper,
3. Name of the journal where it was published,
4. Publication date, and
5. Abstract of the paper (if any).


```{r counts}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website,"/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts,  "[[:digit:],]+")
```

So, I am able to find 2329 papers in total.

```{r get ids}
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(db = 'pubmed',
                term = 'sars-cov-2 trial vaccine',
                  retypr = "abstract",
                    retmax = '250')
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)

# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
# stringr::str_remove_all(ids, "</?Id>")
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")
```

```{r publication}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/",
  path  = "entrez/eutils/efetch.fcgi",
  query = list(
    db = "pubmed",
    id = I(paste(ids, collapse=",")),
    retmax = 250,
    rettype = "abstract"
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```

```{r dataset}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>")
abstracts <- str_replace_all(abstracts, "[[:space:]]+", " ")
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
journal <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
journal <- str_remove_all(journal, "</?[[:alnum:]]+>")
date <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")
date <- str_remove_all(date, "</?[[:alnum:]]+>")
date <- str_replace_all(date, "\\s+", " ")
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Journal = journal,
  Publication_date = date,
  Abstract = abstracts
)
knitr::kable(database[1:20,], caption = "Some papers about Covid19 trail vaccine")
```

## Text Mining
A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.

1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

2.Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

3.Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r download}
dataset <- read.csv("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv")
```


### Question 1
```{r covid_top_10}
dataset %>%
  filter(term == "covid") %>%
  unnest_tokens(token, abstract, token = "words") %>%
  count(token, sort = TRUE) %>%
  top_n(10,n) %>%
  ggplot(aes(n, fct_reorder(token, n)))+
  geom_col()+
  labs(title = "Top 10 Most Frequent Words in Abstracts of Term 'covid'")
```

```{r meningitis_top_10}
dataset %>%
  filter(term == "meningitis") %>%
  unnest_tokens(token, abstract, token = "words") %>%
  count(token, sort = TRUE) %>%
  top_n(10,n) %>%
  ggplot(aes(n, fct_reorder(token, n)))+
  geom_col()+
  labs(title = "Top 10 Most Frequent Words in Abstracts of Term 'meningitis'")
```

```{r prostate_cancer_top_10}
dataset %>%
  filter(term == "prostate cancer") %>%
  unnest_tokens(token, abstract, token = "words") %>%
  count(token, sort = TRUE) %>%
  top_n(10,n) %>%
  ggplot(aes(n, fct_reorder(token, n)))+
  geom_col()+
  labs(title = "Top 10 Most Frequent Words in Abstracts of Term 'prostate cancer'")
```

```{r cystic_fibrosis_top_10}
dataset %>%
  filter(term == "cystic fibrosis") %>%
  unnest_tokens(token, abstract, token = "words") %>%
  count(token, sort = TRUE) %>%
  top_n(10,n) %>%
  ggplot(aes(n, fct_reorder(token, n)))+
  geom_col()+
  labs(title = "Top 10 Most Frequent Words in Abstracts of Term 'cystic fibrosis'")
```

```{r preeclampsia_top_10}
dataset %>%
  filter(term == "preeclampsia") %>%
  unnest_tokens(token, abstract, token = "words") %>%
  count(token, sort = TRUE) %>%
  top_n(10,n) %>%
  ggplot(aes(n, fct_reorder(token, n)))+
  geom_col()+
  labs(title = "Top 10 Most Frequent Words in Abstracts of Term 'preeclampsia'")
```

For top ten words in each search term, most are link words/stop words and search term themselves, and for meningistis, even itself is not in top 10, such results are meaningless for us to analyze.

```{r covid_top_10_nostop}
dataset %>%
  unnest_tokens(token, abstract) %>%
  anti_join(stop_words, by = c("token" = "word")) %>% 
  group_by(term)%>%
  count(token)%>%
  top_n(5, n) %>% 
  knitr::kable(caption ="Top 5 Most Frequent Words in Abstracts after Removing Stop Words")
```

Removing all stop words change what tokens appear as the most frequent.

For term "covid", the top 5 words are "covid" "19""patients""disease""pandemic"

For term "meningitis", the top 5 words are "patients" "meningistis""meningeal""csf""clinical"

For term "prostate cancer", the top 5 words are "cancer" "prostate""patients""treatment""disease"

For term "cystic fibrosis", the top 5 words are "fibrosis""cystic""cf" "patients""disease"

For term "preeclampsia", the top 5 words are "pre""eclampsia""preeclampsia" "women""pregnancy"

### Question 2
```{r bigrams}
dataset %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(10,n) %>%
  ggplot(aes(n, fct_reorder(ngram, n)))+
  geom_col()+
  labs(title = "10 most common bigram")
```

```{r bigram_nostop}
bigrams <- dataset %>%
  unnest_ngrams(ngram, abstract, n = 2) %>% 
  separate(col=ngram, into=c("word1", "word2"), sep = " ") %>%
  select(word1, word2) %>%
  anti_join(stop_words, by = c("word1" = "word")) %>%
  anti_join(stop_words, by = c("word2" = "word")) %>%
  count(word1, word2, sort=TRUE) %>%
  top_n(10, n)
  unite(bigrams, "ngram", c("word1", "word2"), sep = " ") %>%
  ggplot(aes(x = n, y = fct_reorder(ngram, n))) +
  geom_col()+
  labs(title = "10 most common bigram after removing stop words")
```

Before removing stop words, the useful information is still few, but after removing, we get to know more about what the dataset is about.

### Question 3
```{r tf_idf}
dataset %>%
  unnest_tokens(token, abstract) %>%
  count(token, term) %>%
  bind_tf_idf(token, term, n) %>%
  group_by(term) %>%
  top_n(5, tf_idf) %>%
  arrange(desc(tf_idf), .by_group = TRUE) %>%
  select(term, token, n, tf_idf, tf, idf) %>% 
  knitr::kable(caption="5 Tokens from each Search Term with Highest TF-IDF Value")
```
For term "covid", the top 5 words with the highest TF-IDF values are "covid" "pandemic""coronavirus""sars""cov"

For term "meningitis", the top 5 words with the highest TF-IDF values are "meningitis""meningeal""pachymeningitis"csf""meninges"

For term "prostate cancer", the top 5 words with the highest TF-IDF values are "prostate" "androgen""psa""prostatectomy""castration"

For term "cystic fibrosis", the top 5 words with the highest TF-IDF values are "cf""fibrosis""cystic""cftr""sweat"

For term "preeclampsia", the top 5 words with the highest TF-IDF values are "eclampsia""preeclampsia""pragnancy" "maternal""gestational"

Compared to the result from question 1, some of words are same, but I think that TF-IDF provides a more meaningful results since there are more terminology and remove some words that often occur in any kind of medical paper such as "patients" "disease" "treatment"
