---
title: "midterm"
author: "NickyNie"
date: "10/20/2021"
output: html_document
---
# Introduction
This dataset is a collaboration between the COVID Tracking Project and the Boston University Center for Antiracist Research, which shows the counts of cases/death of COVID-19 catrgotized by races for each states in United States. The data was collected from April 12,2020 to March 07,2021. 

## Question: Is death from COVID-19 associated with races?
### Are the trends of incidence/death of different races same across the year in the country level?
### Is the trend in CA similar to that in the country level?
### What are top 5 states with highest death rate from COVID-19 for different races?
### Which race with highest death rate in each state?

# Method
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(tidytext)
library(tidyverse)
library(tidyr)
library(ggforce)
```

## Download data
I download data in csv form from The COVID Tracking Project https://covidtracking.com/race
```{r download}
race <- fread("CRDT.csv")
```

## Preprocessing data

```{r look at NAs}
tab <- race[, .(
NA_total = sum(is.na(race$Cases_Total)),
NA_white = sum(is.na(race$Cases_White)),
NA_black = sum(is.na(race$Cases_Black)),
NA_latinx = sum(is.na(race$Cases_Latinx)),
NA_asian = sum(is.na(race$Cases_Asian)),
NA_AIAN = sum(is.na(race$Cases_AIAN)),
NA_NHPI = sum(is.na(race$Cases_NHPI)),
NA_multi = sum(is.na(race$Cases_Multiracial))
)]
knitr::kable(tab)
```

Based on this table, I decided to take a close look at data of Total, White, Black and Asian since races like Latinx, AIAN, NHPI and multiracial have almost half of missing data, which is not suitable for further analysis. And for the three races categories left, I dropped all the NAs for further analysis. 

I firstly created two new datasets, one is based on state(race_state) and another(race_country) is based on country. Then I calculated death rates based on races of each dataset. I also created a dataset of CA to see if pattern in CA is similar to the pattern of the whole country. And since the number of cases and deaths are cumulative, I chose the latest data(2021.03.07) and created a new dataset called "latest" to find the race with highest death rate in each state.


```{r preprocess, warning=FALSE}
# pick up interested variables
race_state <- unique(race[,.(Date,State,Cases_Total,Cases_White,Cases_Black,Cases_Asian,Deaths_Total,Deaths_White,Deaths_Black,Deaths_Asian)])
# since date here shows to be integer, I changed them into str
race_state$Date <- as.character(race_state$Date)

# drop NAs
race_state <- race_state[!is.na(Cases_Total)]
race_state <- race_state[!is.na(Cases_White)]
race_state <- race_state[!is.na(Cases_Black)]
race_state <- race_state[!is.na(Cases_Asian)]
race_state <- race_state[!is.na(Deaths_Total)]
race_state <- race_state[!is.na(Deaths_White)]
race_state <- race_state[!is.na(Deaths_Black)]
race_state <- race_state[!is.na(Deaths_Asian)]
sum(is.na(race_state))

# Calculate total cases of different races based on date
race_country <- race_state[,.(
  Case_total  = sum(Cases_Total),
  Case_white  = sum(Cases_White),
  Case_black  = sum(Cases_Black),
  Case_asian  = sum(Cases_Asian),
  Death_total = sum(Deaths_Total),
  Death_white = sum(Deaths_White),
  Death_black = sum(Deaths_Black),
  Death_asian = sum(Deaths_Asian)
), by = Date]

# Calculate death rate
race_state  [, DR_total := Deaths_Total/Cases_Total]
race_state  [, DR_white := Deaths_White/Cases_White]
race_state  [, DR_black := Deaths_Black/Cases_Black]
race_state  [, DR_asian := Deaths_Asian/Cases_Asian]
race_country[, DR_total := Death_total/Case_total]
race_country[, DR_white := Death_white/Case_white]
race_country[, DR_black := Death_black/Case_black]
race_country[, DR_asian := Death_asian/Case_asian]

# Stick to CA data
CA <- filter(race_state, State == "CA")

# Race with highest DR in each state.
latest <- filter(race_state, Date == "20210307")
latest <- latest[, highest_DR := 
         fifelse(DR_white == max(DR_white,DR_black,DR_asian),"White",
          fifelse(DR_black == max(DR_white,DR_black,DR_asian),"Black","Asian"
          )), by = State]
```

# Preliminary Results
```{r country}
ggplot(data = race_country) + 
  geom_point(mapping = aes(x = Date, y = DR_white, color = "DR_white"))+
  geom_point(mapping = aes(x = Date, y = DR_black, color = "DR_black"))+
  geom_point(mapping = aes(x = Date, y = DR_asian, color = "DR_asian"))+
  geom_point(mapping = aes(x = Date, y = DR_total,color = "DR_total"))+
  labs(x = "Date", y = "Death Rate")+
  labs(title = "COVID-19 Death rate of different races in United States")+
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=1))
```

From this graph, we can see that the trend of all DR is gradually decreasing, the total death rate is always the lowest one since there are a lot of cases of other or unknown races which largely increase the denominator. And although death rate of white is higher than the other two races initially, they approach to the same level finally with a little bit of difference.

```{r CA}
ggplot(data = CA) + 
  geom_point(mapping = aes(x = Date, y = DR_white, color = "DR_white"))+
  geom_point(mapping = aes(x = Date, y = DR_black, color = "DR_black"))+
  geom_point(mapping = aes(x = Date, y = DR_asian, color = "DR_asian"))+
  geom_point(mapping = aes(x = Date, y = DR_total,color = "DR_total"))+
  labs(x = "Date", y = "Death Rate")+
  labs(title = "COVID-19 Death rate of different races in California")+
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=1))
```

The general trend is similar to that of country but it seems that four lines go up slightly at last. But we can still see a trend of DR of three races approaching to the same level at the final stage of the data.

```{r top5}
top5_total <- latest[order(DR_total,decreasing = TRUE)][1:5]
top5_total
top5_white <- latest[order(DR_white,decreasing = TRUE)][1:5]
top5_white
top5_black <- latest[order(DR_black,decreasing = TRUE)][1:5]
top5_black
top5_asian <- latest[order(DR_asian,decreasing = TRUE)][1:5]
top5_asian
```

The top5 states with highest death rate of total are  `r top5_total$State`

The top5 states with highest death rate of white are  `r top5_white$State`

The top5 states with highest death rate of black are  `r top5_black$State`

The top5 states with highest death rate of asian are  `r top5_asian$State`
