---
title: "Assignment_02"
author: "NickyNie"
date: "10/4/2021"
output: 
  - html_document
  - github_document
always_allow_html: true
---

```{r setup, include=FALSE, cache = TRUE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(dplyr)
library(leaflet)
library(ggforce)
```

# Data Wrangling
## Download data
```{r download}
ind <- fread("chs_individual.csv")
reg <- fread("chs_regional.csv")
```

## Merge datasets
```{r merge}
data <- merge(
  x = ind,
  y = reg,
  by ="townname",
  all.x = TRUE,
  all.y = FALSE
)
```

## Checking duplications
```{r row_nuumber}
dim(ind)
dim(reg)
dim(data)
```
No duplication

## Impute NAs
The key variables are bmi, fev, asthma, smoke pm2.5 mass and gasstove, check for NAs
```{r check_NAs}
sum(is.na(data$bmi))
sum(is.na(data$fev))
sum(is.na(data$smoke))
sum(is.na(data$gasstove))
sum(is.na(data$asthma))
sum(is.na(data$pm25_mass))
```

```{r impute}
data[, m_bmi := mean(bmi, na.rm = TRUE), by = .(male, hispanic)]
data[, bmi := ifelse(is.na(bmi), m_bmi, bmi)]
data[, m_fev := mean(fev, na.rm = TRUE), by = .(male, hispanic)]
data[, fev := ifelse(is.na(fev), m_fev, fev)]
data[, m_smoke := mean(smoke, na.rm = TRUE), by = .(male, hispanic)]
data[, smoke := ifelse(is.na(smoke), m_smoke, smoke)]
data[, m_gas := mean(gasstove, na.rm = TRUE), by = .(male, hispanic)]
data[, gasstove := ifelse(is.na(gasstove), m_gas, gasstove)]
data[, m_asthma := mean(asthma, na.rm = TRUE), by = .(male, hispanic)]
data[, asthma := ifelse(is.na(asthma), m_asthma, asthma)]
sum(is.na(data$bmi))
sum(is.na(data$fev))
sum(is.na(data$smoke))
sum(is.na(data$gasstove))
sum(is.na(data$asthma))
```

## Create obesity level variable using BMI measurement
```{r BMI}
data <- data[, obesity_level := fifelse(bmi<14, "Underweight",
                                    fifelse(bmi >= 14  & bmi < 22, "Normal",
                                        fifelse(bmi >= 22 & bmi < 24,"Overweight","Obesity")))]
```

knit a table
```{r table}
tab <- data[, .(
  Max_BMI = max(bmi),
  Min_BMI = min(bmi),
  Total_Number = .N
), by = obesity_level]

knitr::kable(tab)
```

## Create smoke-gas-exposure variable
```{r smoke_gas_exposure}
data <- data[, smoke_gas_exposure := fifelse(smoke == 0 & gasstove == 0, "No_Exposure",
                                          fifelse(smoke == 1 & gasstove == 0, "Smoke_Exposure",
                                              fifelse(smoke == 0 & gasstove == 1, "Gas_Exposure","Both_Exposure")))]
```

table of town
```{r}
tab <- data[, .(
  avg_fev = mean(fev),
  sd_fev = sd(fev),
  avg_asthma = mean(asthma),
  sd_asthma = sd(asthma)
), by = townname]

knitr::kable(tab)
```

table of sex
```{r}
tab <- data[, .(
  avg_fev = mean(fev),
  sd_fev = sd(fev),
  avg_asthma = mean(asthma),
  sd_asthma = sd(asthma)
), by = male]

knitr::kable(tab)
```

table of obesity level
```{r}
tab <- data[, .(
  avg_fev = mean(fev),
  sd_fev = sd(fev),
  avg_asthma = mean(asthma),
  sd_asthma = sd(asthma)
), by = obesity_level]

knitr::kable(tab)
```

table of smoke_gas_exposure
```{r}
tab <- data[, .(
  avg_fev = mean(fev),
  sd_fev = sd(fev),
  avg_asthma = mean(asthma),
  sd_asthma = sd(asthma)
), by = smoke_gas_exposure]

knitr::kable(tab)
```

# Looking at the data
check for dimension
```{r dim}
dim(data)
```

headers and footers
```{r headers}
head(data)
```



```{r footer}
tail(data)
```

variable
```{r variable}
str(data)
```
key variable
```{r key}
summary(data$fev)
summary(data$bmi)
summary(data$smoke_gas_exposure)
summary(data$pm25_mass)
```

### Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r scatterplots}
ggplot(
  data[!is.na(bmi)],
  mapping = aes(x = bmi, y = fev, color = townname))+
  geom_point()+
  geom_smooth(method = "lm", color = "black")+
  facet_wrap(~townname)+
  labs(title = "Scatterplots of BMI vs FEV by townname", x = "BMI", y = "FEV")
```

From the graph, we can see that bmi has a positive association with fev for all towns. The change in fev is greater in Upland when bmi changes and least in Alpine.

### Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r his_by_bmi}
ggplot(
  data[!is.na(obesity_level) & !is.na(fev)],
  mapping = aes(x = fev, fill = obesity_level))+
  geom_histogram()+
  labs(title = "Histogram of FEV by BMI category", x = "FEV")+
  scale_fill_brewer(palette = "RdPu")
```
It seems that for the category of Normal, it follows a normal distribution and no clear distribution for other categories.

```{r his_by_exp}
ggplot(
  data[!is.na(smoke_gas_exposure) & !is.na(fev)],
  mapping = aes(x = fev, fill = smoke_gas_exposure))+
  geom_histogram()+
  labs(title = "Histogram of FEV by smoke-gas-exposure", x = "FEV")+
  scale_fill_brewer(palette = "GnBu")
```
In this graph, it seems that all the the distribution for all the categories is normal distribution.

### Barchart of BMI by smoke/gas exposure.
```{r bar}
ggplot(
  data[!is.na(smoke_gas_exposure) & !is.na(obesity_level)],
  mapping = aes(x = obesity_level, fill = smoke_gas_exposure))+
  geom_bar()+
  labs(title = "Barchart of BMI by smoke-gas-exposure", x = "BMI_category")+
  scale_fill_brewer(palette = "Reds")
```

For all bmi categories, the order of number of counts from greater to smaller is both_exposure, gas_exposure, no_exposure and smoke_exposure.

### Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r sum_bmi}
ggplot(
  data[!is.na(fev) & !is.na(obesity_level)],
  mapping = aes(x = obesity_level, y = fev))+
  stat_summary(fun.min = min, fun.max = max, fun = median)+
  labs(title = "statistical summary of FEV by BMI")
```

The category of normal has a relatively small median but with greatest range. And for category of overweight, it has a relatively great median but with a small range.

```{r sum_exp}
ggplot(
  data[!is.na(fev) & !is.na(smoke_gas_exposure)],
  mapping = aes(x = smoke_gas_exposure, y = fev))+
  stat_summary(fun.min = min, fun.max = max, fun = median)+
  labs(title = "statistical summary of FEV by smoke-gas-exposure")
```

It seems that based on exposure categories, the medians are almost same for all of them.

### A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

Choose a visualization to examine whether PM2.5 mass is associated with FEV.
